Brazilian Consumer Complaints by Luiz Gerosa
========================================================

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(knitr)
library(ggplot2)
library(dplyr)
library(tidyr)
library(stringr)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data
Sys.setlocale(category = "LC_ALL", locale = "pt_BR.UTF-8")
ds = read.csv("data/reclamacoes-fundamentadas-sindec-2015.csv", 
              sep=';', encoding='utf-8', stringsAsFactors = FALSE, strip.white = TRUE)

# keeping only columns used in this analysis
ds <- subset(ds, select=c(AnoCalendario, DataArquivamento, DataAbertura, UF,
                          strRazaoSocial, Tipo,
                          Atendida, DescricaoProblema, SexoConsumidor,
                          FaixaEtariaConsumidor))

# renaming
names(ds) <- c('year', 'closingDate', 'registerDate', 'state', 'companyName',
               'type', 'resolved', 'issue', 'gender', 'age')

# replacing empty cells with NA
ds[ds == ''] <- NA
ds[ds == 'NULL'] <- NA
ds$gender[ds$gender == 'N'] <- NA

# removing rows with any missing value (NA)
ds <- na.omit(ds)

# keeping only complaints of companies (type == 1)
ds <- ds[ds$type == 1,]
ds$type <- NULL

# fixing data types
ds$closingDate <- as.Date(ds$closingDate)
ds$registerDate <- as.Date(ds$registerDate)
ds$state <- as.factor(ds$state)
ds$companyName <- as.factor(ds$companyName)
ds$resolved <- as.factor(ds$resolved)
ds$gender <- as.factor(ds$gender)
ds$age <- as.factor(ds$age)
ds$issue <- as.factor(ds$issue)

# translating age ranges
ds$age <- recode(ds$age,
                 `até 20 anos` = '<20',
                 `entre 21 a 30 anos` = '21 to 30',
                 `entre 31 a 40 anos` = '31 to 40',
                 `entre 41 a 50 anos` = '41 to 50',
                 `entre 51 a 60 anos` = '51 to 60',
                 `entre 61 a 70 anos` = '61 to 70',
                 `mais de 70 anos` = '>70',
                 `Nao Informada` = 'not informed')

# translating S (sim) to Y (yes)
ds$resolved <- recode(ds$resolved, S = 'Y', N = 'N')

# creating a new continuous variable
ds$elapsedDays <- as.numeric(ds$closingDate - ds$registerDate)

# cleaning wrong chars
ds$companyName <- sub('[@|\\.]$', '', ds$companyName)

# very basic normalizations of company types
ds$companyName <- sub('S[\\.|\\/]+A.*$', 'S/A', ds$companyName)
ds$companyName <- sub('LTDA*$', 'LTDA', ds$companyName)

# fix typos for most frequent companies
ds$companyName <- sub('SERVICOS', 'SERVIÇOS', ds$companyName)
ds$companyName <- sub('ECONOMICA', 'ECONÔMICA', ds$companyName)
ds$companyName <- sub('COMUNICA[C|Ç][A|Ã]O', 'COMUNICAÇÃO', ds$companyName)
ds$companyName <- sub('COMUNICACOES', 'COMUNICAÇÕES', ds$companyName)
ds$companyName <- sub('- CEF$', '', ds$companyName)
ds$companyName <- sub('^OI.+S/A$', 'OI S/A', ds$companyName)

ds$companyName <- str_trim(ds$companyName)

```

# Introduction
When Brazilian consumers need to resolve a dispute with business the first step 
is to go to a local Procon (Consumer Protection Agency) and file a complaint.
The Procon assists the consumer and intermediates the resolution with the
company

# DataSet Overview

The [dataset](http://dados.gov.br/dataset/cadastro-nacional-de-reclamacoes-fundamentadas-procons-sindec1)
used in this analysis contains information about complaints made in 2015 and it
was download from official Brazilian government open data website: http://dados.gov.br. 


```{r echo=FALSE, results = 'asis'}
kable(head(ds))
```

### Summary
```{r echo=FALSE}
summary(ds)
```

### Top 10 most common complaints (in Portugues)
```{r echo=FALSE}
issueFrequency <- count(ds, issue)
issueFrequency <- arrange(issueFrequency, desc(n))
kable(head(issueFrequency, 10))
```

### Top 10 most frequent companies
```{r echo=FALSE}
top_companies <- count(ds, companyName)
top_companies <- arrange(top_companies, desc(n))
top_companies <- top_companies[1:100, ]
kable(head(top_companies, 10))
```

# Univariate Plots Section

```{r echo=FALSE}
resolved <- ds %>%
  group_by(resolved) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))

resolutionRate <- as.double(resolved[2, 'freq'])

pct <- paste(round(resolved$freq*100), '%', sep = '')
pie(resolved$freq, paste(resolved$resolved, pct, sep = ' - '))
```

The resolution rate is about 65%.

```{r echo=FALSE}
gender <- ds %>%
  group_by(gender) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))

pct <- paste(round(gender$freq*100), '%', sep='')
pie(gender$freq, paste(gender$gender, pct, sep = ' - '))
```

54% of complaints came from women vs 46% from men.

```{r echo=FALSE}
qplot(x = ds$state, data = ds)
```

The state with most complaints is São Paulo.

```{r echo=FALSE}
qplot(x = ds$age, data = ds)
```

Most complaints are made by people between 31 and 40 years old.

```{r echo=FALSE, warning=FALSE}
ggplot(aes(x = elapsedDays), data = ds) +
  geom_histogram(binwidth = 5) +
  scale_x_continuous(limits = c(0, 500), breaks = seq(0, 500, 20))
```

The number of days to close an issue has a long tail.

# Univariate Analysis

### What is the structure of your dataset?

After cleaning missing values, removing columns not used in this analysis and 
adding a new one, the dataset contains around 221k complaints and 10 variables.

```{r}
dim(ds)
```

### What is/are the main feature(s) of interest in your dataset?

The main feature is _"resolved"_ which indicates if the complaint was resolved
with the company.

### What other features in the dataset do you think will help support your \
investigation into your feature(s) of interest?

Categorical features like gender, state, age, etc.

### Did you create any new variables from existing variables in the dataset?

Yes. I calculated the elapsed time in days between the registration date and the
closing date.

### Of the features you investigated, were there any unusual distributions? \
Did you perform any operations on the data to tidy, adjust, or change the form \
of the data? If so, why did you do this?

I did some basic data wrangling in the company name so it can be counted correctly.

# Bivariate Plots Section

```{r echo=FALSE, Bivariate_Plots}

gender_resolved <- ds %>%
  group_by(gender, resolved) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n))

ggplot(gender_resolved, aes(x=gender,y=freq,fill=factor(resolved)))+
  geom_bar(stat="identity") +
  geom_hline(yintercept = resolutionRate)
```

Men and women have practically the same chance to have their issues resolved.

```{r echo=FALSE}

state_resolved <- ds %>%
  group_by(state, resolved) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  arrange(freq)

ggplot(state_resolved, aes(x=state, y=freq, fill=factor(resolved))) +
  geom_bar(stat="identity") +
  geom_hline(yintercept = resolutionRate)
```

Consumers in some states have a much higher chance to resolve their issues.

```{r echo=FALSE}

age_resolved <- ds %>%
  group_by(age, resolved) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  arrange(freq)

ggplot(age_resolved, aes(x=age, y=freq, fill=factor(resolved))) +
  geom_bar(stat="identity") +
  geom_hline(yintercept = resolutionRate)
```

All age intervals have practically the same resolution rate.
When the age is not informed, the resolution rate is smaller.

```{r echo=FALSE}

company_resolved <- ds %>%
  group_by(companyName, resolved) %>%
  summarise (n = n()) %>%
  spread(resolved, n, fill=0) %>%
  arrange(desc(N+Y)) %>%
  mutate(freq = Y / (N+Y))

kable(head(company_resolved, 10))
```

Resolution rate for the top 10 most frequent companies. OI S/A is the company
with most complaints but also has a very high resolution rate.

```{r echo=FALSE, warning=FALSE}
ggplot(ds, aes(x = elapsedDays, color=resolved)) +  
  geom_freqpoly(binwidth=10) +
  scale_x_continuous(limits = c(0, 500), breaks = seq(0, 500, 20))
```

The shape of the distribution of the number of days to close an issue doesn't
change if the issue was resolved or not.

```{r echo=FALSE, warning=FALSE}
ggplot(ds, aes(x = elapsedDays, color=resolved)) +  
  geom_density() +
  scale_x_continuous(limits = c(0, 200), breaks = seq(0, 200, 25))
```

The issue has a better chance to be resolved between 25 and 60 days after its
registration date.

```{r echo=FALSE}

state_gender <- ds %>%
  group_by(state, gender) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  arrange(freq)

man_rate = as.double(gender[2, 'freq'])

ggplot(state_gender, aes(x=state, y=freq, fill=factor(gender))) +
  geom_bar(stat="identity") +
  geom_hline(yintercept = man_rate)
```

Women and men complaints per state are uniformly distributed.

```{r echo=FALSE}

state_resolved_elapsedays <- ds %>%
  group_by(state, resolved) %>%
  summarise(median = median(elapsedDays))

ggplot(state_resolved_elapsedays, aes(x=state, y=median)) +
  geom_bar(aes(fill = resolved), position = "dodge", stat="identity")
```

Some states (e.g. AP, PB and TO) have a huge difference between 
the median time to close an issue when it was resolved or not.

```{r echo=FALSE}

ggplot(subset(state_resolved_elapsedays, (resolved == 'Y')), aes(x=state, y=median)) +
  geom_bar(aes(fill = resolved), position = "dodge", stat="identity")
```

Observing only resolved issues, we can have a better sense of the median time 
to resolve an issue on different states.

```{r echo=FALSE}

age_gender <- ds %>%
  group_by(age, gender) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  arrange(freq)

ggplot(age_gender, aes(x=age, y=freq, fill=factor(gender))) +
  geom_bar(stat="identity") +
  geom_hline(yintercept = man_rate)
```

Women and men complaints per age interval are uniformly distributed.
Men do not inform their age more frequently than women.

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

The resolution rate varies a lot depending on the state and the number of days
it takes to close the complaint.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

The median time to resolve an issue varies a lot state by state. 

### What was the strongest relationship you found?

The state where the complaint was filed has a strong impact on the resolution
rate.

------

# Final Plots and Summary

### Resolution rate per state
```{r echo=FALSE, Plot_One}

state_resolved <- subset(state_resolved, (resolved == 'Y'))

state_resolved$state <- factor(state_resolved$state,
                          levels(reorder(state_resolved$state, state_resolved$freq)))

breaks = c(resolutionRate, seq(0, 1, 0.25), max(state_resolved$freq))

ggplot(state_resolved, aes(x=state, y=freq)) +
  geom_bar(stat="identity", fill='#00BFC4') +
  annotate("text", x = 3.1, y = resolutionRate + 0.03, label = "Avg. resolution rate") +
  geom_hline(yintercept = resolutionRate, linetype="dashed") +
  scale_y_continuous(labels = scales::percent, breaks = breaks) +
  xlab('State') + 
  ylab('Resolution rate') +
  ggtitle("Resolution rate per state")

```

The state where the complaint was filed has a big impact on the chance to it be
resolved. The difference between the best resolution rate (in PB) and the worst
(in PR) is 33%.

### Median interval to resolve an issue per state
```{r echo=FALSE, Plot_Two}

state_resolved_elapsedays <- subset(state_resolved_elapsedays, (resolved == 'Y'))
medianElapsedDaysResolved <- median(ds[resolved == 'Y', 'elapsedDays'])

state_ordered <- reorder(state_resolved_elapsedays$state, state_resolved_elapsedays$median)
state_resolved_elapsedays$state <- factor(state_resolved_elapsedays$state, levels(state_ordered))

breaks <- sort(c(medianElapsedDaysResolved, seq(0, 600, 150)))

ggplot(state_resolved_elapsedays, aes(x=state, y=median)) +
  geom_bar(stat="identity", fill = '#00BFC4') +
  annotate("text", x = 2, y = medianElapsedDaysResolved +20, label = "Median") +
  geom_hline(yintercept = medianElapsedDaysResolved, linetype="dashed") +
  scale_y_continuous(breaks = breaks) +
  xlab('State') + 
  ylab('Days') +
  ggtitle("Median interval to resolve an issue per state") 
```

Different states vary a lot the interval between the registration and close date.
The chart above shows the median interval to resolve an issue (resolved = 'Y')
per state.

### Plot Three
```{r echo=FALSE, Plot_Three, warning=FALSE}

breaks = sort(c(24, 64, seq(100, 250, 50)))

ggplot(ds, aes(x = elapsedDays)) +  
  geom_density(aes(color = resolved), alpha=0.4) +
  scale_x_continuous(limits = c(0, 250), breaks = breaks) +
  geom_vline(xintercept=24, linetype="dashed") +
  geom_vline(xintercept=64, linetype="dashed") +
  xlab('Elapsed days between regisration and close date') + 
  ggtitle("Density curve of the elapsed days interval")
```

The chart above shows that the interval between 24 and 64 days has a greater
density of resolved issues, i.e. the resolution rate is greater,
as can be observed in the table below:

```{r echo=FALSE, Plot_Three_Aux_Table, warning=FALSE}
ds$elapsedDaysIntervals <- cut( ds$elapsedDays, 
                                breaks = c(0, 25, 64, 10^4),
                                include.lowest = TRUE)

elapseDaysIntervals_resolved <- ds %>%
  group_by(elapsedDaysIntervals, resolved) %>%
  summarise (n = n()) %>%
  mutate(freq = n / sum(n)) %>%
  arrange(desc(resolved), elapsedDaysIntervals)

kable(head(elapseDaysIntervals_resolved[elapseDaysIntervals_resolved$resolved == 'Y', c(1, 4)]))
```

This information will help to build a predictive model in a future work.

------

# Reflection

This exploratory analysis gave an overview of the performance of Procons in 
different states. For example, in Brasilia, capital of Brazil, the Procon has
a very poor performance in both the resolution rate (second worst) and
the time it took to resolve an issue (almost 2 years).

Ideas for the future:

* Use data from other years to analyse trends in data.
* Come up with an indicator of the quality of the Procon considering the resolution
rate and the time it takes to resolve the issue.
* Come up with an indicator for the companies.
* Come up with a predictive model to give the probability of an issue be resolved.
